---
title: "Homework 3 - FE 570"
author: "Siddharth Iyer"
date: "3/30/2022"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages
```{r, warning=FALSE, message=FALSE}
library(xts)
library(highfrequency)
```

Load Dataset
```{r}
# load the TAQ dataset
Sys.setenv(TZ = "GMT") # work in East Coast Time Zone
options(digits.secs=3)
load("sampleTQdata.RData")
```

Problem 3.1.1: Calibrate Roll Model 
```{r, warning=FALSE, message=FALSE}
# Calibrate the Roll model on the tqdata
pr <- as.numeric(tqdata$PRICE)
dpr <- diff(pr)  # Delta price
acpr <- acf(dpr, lag.max=20, type="correlation", 
            plot=TRUE, main="Autocorrelation")

# acf: both autocorrelation and autocovariance

plot(acpr, col="red", lag.max=20, 
     main="Autocorrelation of price changes")

# Roll estimate of bid-ask spread

covpr <- acf(dpr, lag.max=20, type="covariance", plot=FALSE)

gamma0 <- sd(dpr)^2
#print(gamma0)

gamma0alt <- covpr$acf[1]
#print(gamma0alt)

gamma1 <- covpr$acf[2]
#print(gamma1)

cparam.1 <- sqrt(-covpr$acf[2])
cat("bid-ask spread (2*c): ", 2*cparam.1, "\n")

sig2u <- gamma0 + 2* gamma1
sigu.1 <- sqrt(sig2u)
cat("sigma: ", sigu.1, "\n")

```

Problem 3.1.2: Calibrate Roll Model with log(pt)
```{r, warning=FALSE, message=FALSE}
# Calibrate the Roll model on the tqdata
pr <- as.numeric(tqdata$PRICE)
log_pr <- log(pr)
dpr <- diff(log_pr)  # Delta price
acpr <- acf(dpr, lag.max=20, type="correlation", 
            plot=TRUE, main="Autocorrelation")

# acf: both autocorrelation and autocovariance

plot(acpr, col="red", lag.max=20, 
     main="Autocorrelation of price changes")

# Roll estimate of bid-ask spread

covpr <- acf(dpr, lag.max=20, type="covariance", plot=FALSE)

gamma0 <- sd(dpr)^2
#print(gamma0)

gamma0alt <- covpr$acf[1]
#print(gamma0alt)

gamma1 <- covpr$acf[2]
#print(gamma1)

cparam.2 <- sqrt(-covpr$acf[2])
cat("bid-ask spread (2*c): ", 2*cparam.2, "\n")

sig2u <- gamma0 + 2* gamma1
sigu.2 <- sqrt(sig2u)
cat("sigma: ", sigu.2, "\n")

```
When we are using log(price) and applying diff on the datasets, we are effectively taking the spread on returns data. I don't this this makes much sense. 


Problem 3.1.3
```{r}
realizedVar <- function(q){
  pr <- as.numeric(tqdata$PRICE)
  rCov(diff(pr, lag=q, differences=1)/q)
}
# compute the signature plot sigma.day(q) = sqrt(RV(q))
sig_data <- NULL

for(q in 1:100){
  sig_data <- c(sig_data, sqrt(realizedVar(q)))
}

plot(sig_data, type ="l", main="Signature plot")
```

We can see that as the sampling interval decreases, the realizedVol goes to 0 because it all cancels out. This makes sense because the volatility is normally distributed around mean, so the with more data being smoothed by the resample, the closer the realizedVol will be to 0.  

This code is taken from the lectures. 


Problem 3.1.4
```{r}
n_trades = length(tqdata)
q_5min = n_trades*5/390
cat("Realized Var: ", q_5min*realizedVar(q_5min), "\n")

sig_roll = sqrt(n_trades * (sigu.1)^2)
cat("Roll: ", sig_roll, "\n")

```

Incredible! The Roll model and the Realized volatility is approximately the same at a log of 5 minutes. 























