---
title: "Coca Cola Market Impact"
output: html_notebook
---

Load the data and filter for trades in ADF exchange only

```{r, warning=FALSE, message=FALSE}
# Libraries
library(xts)
library(highfrequency)
library(quantmod)
library(stats)


Sys.setenv(TZ = "EST")  # dataset is timestamped in GMT
options(digits.secs=3)

# Load Data
load("taqdata_KO_20220301EX.RData")
options(digits.secs=12)
tqdata_zoo = fortify.zoo(tqdata)
write.table(tqdata_zoo, "tqdata_new.txt", row.names = TRUE)

# Filter trades only in Exchange: ADF
tqdata = tqdata[tqdata$EX == "ADF"]
head(tqdata)
```

Filter for trades between 10am - 3pm.

```{r, message=FALSE, warning=FALSE}
# keep only a subset of trading hours (2022-03-01 10am - 3pm)
sampledata <- '2022-03-01 10:00:00::2022-03-01 15:00:00'
tqdata1hr <- tqdata[sampledata]

length(tqdata$PRICE) # 11,127 trades
```

Calculate trade signs & signed trade volumes. Calculate bids, mids, changes in mid prices & trade prices

```{r, message=FALSE, warning=FALSE}
# calculate trade signs & signed trade volumes
tradeSigns <- getTradeDirection(tqdata1hr)
signed.x <- as.numeric(tqdata1hr$SIZE)*tradeSigns
signed.x <- signed.x[1:(length(signed.x)-1)] # trim to same size as dm

# calculate bids, mids, changes in mid prices & trade prices
prices <- as.numeric(tqdata1hr$PRICE)
asks <- as.numeric(tqdata1hr$OFR)
bids <- as.numeric(tqdata1hr$BID)
mids <- 0.5*bids + 0.5*asks
dm <- diff(mids)
dp <- diff(prices)
```

```{r, message=FALSE, warning=FALSE}
plot(prices, type="l", col="blue", main = "Bid and Ask Prices - No Jumps in Window")
lines(mids, type="l", col="red")
```
The ACF plot shows that Trade Signs show significant autocorrelation, while Signed Trade Sizes show no such autocorrelation. 

```{r, message=FALSE, warning=FALSE}
acf(tradeSigns)
acf(signed.x)
```
Price Impact Chart

```{r, warning=FALSE, message=FALSE}
plot(signed.x, dm, pch=20,col="blue",
     main="mid-price change vs trade size", 
     xlab="Signed size", ylab="d(mid)", xlim = c(-5000, 5000), ylim=c(-0.1,0.1))

```

Here we try to determine whether 0/1-lagged signed trade sizes is enough to determine price impact. CCF analysis shows must significance at lag 0 and 1. 

```{r, warning=FALSE, message=FALSE}
# determine optimal lag from ccf analysis
library(stats)

dm.ts <- ts(dm)
signed.x.ts <- ts(signed.x)

# cov(dm,signed.x) =  0.4723287
# cor(dm,signed.x) = 0.06152167

# we expect that dm lags signed_x
dm.x.ccf <- ccf(signed.x.ts, dm.ts)
```
Glosten-Harris Model fit to Data

```{r, warning=FALSE, message=FALSE}
######################################################
# calibrate the Glosten-Harris model (Roll + price impact)
# p(t) = m(t) + c*d(t) + lambda*x(t)
# m(t) = m(t-1) + lambda*x(t-1) + u(t)
####################################################

# Find Lambda 
# m(t) = m(t-1) + lambda*x(t-1) + u(t)
n <- length(signed.x)
dm.lag <- dm

(fitGlHarris <- lm(dm.lag ~ signed.x, subset=(abs(signed.x)>100)&abs(signed.x)<10000))
lambda <- fitGlHarris$coeff[2]

plot(signed.x, dm, pch=20,col="blue",
     main="mid-price change vs trade size (ADF 10am-3pm)", 
     xlab="Signed size", ylab="d(mid)", xlim=c(-5000,5000), ylim=c(-0.05,0.05))

abline(fitGlHarris, col="red")

# lambda = 4.666e-07

# Find c
# p(t) = m(t) + c*d(t) + lambda*x(t)
ns <- length(signed.x)
priceDiff <- prices[1:ns] - mids[1:ns] - lambda*signed.x

(fitGlHarris2 <- lm(priceDiff ~ tradeSigns[1:ns]))

summary(fitGlHarris2)
summary(fitGlHarris)
# c = -1.946e-05

#s = 2*(c + lambda*|x|)
#cat("Bid-Ask Spread = ", 2*(c + lambda * abs()))

```
Foucault Approach to measuring price impact

```{r, message=FALSE, warning=FALSE}
################################################
# Price-impact regression using Eq.(5.4) in Foucault, Pagano, Roell
# dp(t) = lambda*d(t) + gamma*d(d(t)) + eps(t)
# regress dp ~ d, dd
# bid-ask spread = 2*(lambda + gamma)

dp <- diff(prices)

dt <- diff(tradeSigns)
tradeSigns_trim <- tradeSigns[1:(length(tradeSigns)-1)] # trim to same size as dt

length(dp)
length(dt)

(fitPriceImpact <- lm(dp ~ tradeSigns_trim + dt))

#Coefficients:
#    (Intercept)  tradeSigns_trim               dt  
#     -2.184e-05        1.089e-03        3.333e-03   

residDp <- dp + 0.0001 - 0.00298*tradeSigns_trim - 0.0077869*dt

plot(residDp, type="p", col="black", ylim=c(-0.1,0.1))
plot(dp, type="p", col="black", ylim=c(-0.1,0.1))

plot(dt, dp, pch=20,col="blue",
     main="price change vs price reversal", 
     xlab="Signed size", ylab="d(mid)", xlim=c(-2,2), ylim=c(-0.5,0.5))

abline(fitPriceImpact, col="red")

cat("Bid-Ask Spread: ", 2*(3.333e-03 + 1.089e-03))
```

```{r, message=FALSE, warning=FALSE}
################################################
# Price-impact regression using Eq.(5.4) in Foucault, Pagano, Roell
# dp(t) = lambda*d(t) + gamma*d(d(t)) + eps(t)
# regress dp ~ d, dd
# bid-ask spread = 2*(lambda + gamma)

# Modified Foucault
# dp(t) = lambda*x(t) + gamma*d(d(t))*dx(t) + eps(t)

trade_sizes = as.numeric(tqdata1hr$SIZE)
dp <- diff(prices)

dt <- diff(tradeSigns)
tradeSigns_trim <- tradeSigns[1:(length(tradeSigns)-1)] # trim to same size as dt
signed.x1 = tradeSigns_trim*trade_sizes[-1]
signed.x2 = dt*trade_sizes[-1]

length(dp)
length(dt)

(fitPriceImpact <- lm(dp ~ signed.x1 + signed.x2))
(fitPriceImpact_mid <- lm(dm ~ signed.x1 + signed.x2 ))

#Coefficients:
#    (Intercept)  tradeSigns_trim               dt  
#     -2.184e-05        1.089e-03        3.333e-03   

plot(dp, type="p", col="black", ylim=c(-0.1,0.1))

plot(signed.x, dp, pch=20,col="blue", main = "Modified Foucault",
     xlim=c(-5000,5000), ylim=c(-0.05,0.05))
abline(fitPriceImpact, col="red")

plot(signed.x, dm, pch=20,col="blue", main = "Modified Foucault",
     xlim=c(-5000,5000), ylim=c(-0.05,0.05))
abline(fitPriceImpact_mid, col = "red")

cat("Bid-Ask Spread: ")

2*(2.386e-07 + 1.911e-07)
```








